---
title: "STA1004 Assignment 3"
author: "Giyu"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

# =====================================================
# PART 1 — BACTERIAL DOUBLING TIME BY POLLUTANT
# =====================================================

```{r bacteria-data}
bacteria_data <- data.frame(
  pollutant = factor(
    c("Diesel","Control","Diesel","Control","Mercury","Pesticide","Pesticide","Diesel","Mercury","Control","Pesticide","Diesel","Pesticide","Cadmium","Mercury","Diesel","Mercury","Cadmium","Cadmium","Control","Pesticide","Control","Cadmium","Diesel","Pesticide","Diesel","Cadmium","Mercury","Pesticide","Control","Diesel","Mercury","Cadmium","Diesel","Mercury","Pesticide","Mercury","Control","Pesticide","Control","Mercury","Pesticide","Control","Cadmium","Diesel","Mercury","Control","Control","Control","Pesticide","Pesticide","Diesel","Mercury","Pesticide","Cadmium","Cadmium","Mercury","Diesel","Diesel","Cadmium","Cadmium","Diesel","Cadmium","Pesticide","Mercury","Pesticide","Cadmium","Cadmium","Cadmium","Cadmium","Cadmium","Cadmium","Control","Diesel","Mercury","Control","Control","Diesel","Mercury","Pesticide","Control","Diesel","Pesticide","Mercury","Pesticide","Pesticide","Diesel","Control","Control","Diesel","Control","Pesticide","Diesel","Control","Cadmium","Diesel","Mercury","Pesticide","Cadmium","Cadmium","Cadmium","Mercury","Diesel","Pesticide","Mercury","Pesticide","Cadmium","Mercury","Mercury","Cadmium","Control","Mercury","Diesel","Control","Pesticide","Cadmium","Pesticide","Control","Diesel","Diesel","Cadmium","Control","Mercury","Cadmium","Pesticide","Diesel","Diesel","Mercury","Pesticide","Cadmium","Control","Diesel","Control","Control","Mercury","Diesel","Mercury","Cadmium","Control","Pesticide","Pesticide","Mercury","Mercury","Control","Pesticide","Mercury","Diesel","Cadmium","Mercury","Control"),
    levels = c("Control","Diesel","Mercury","Pesticide","Cadmium")
  ),
  doubling_time = c(37,28,33,38,39,35,36,36,37,33,32,33,35,29,36,40,31,34,26,35,33,36,30,35,38,35,34,39,44,41,34,31,35,24,28,34,36,44,33,35,32,35,35,34,26,32,34,36,37,37,33,29,34,37,32,33,31,35,42,29,37,26,35,23,26,34,27,30,32,32,29,43,28,34,35,32,35,25,33,34,34,36,37,29,27,30,35,34,33,42,36,30,37,32,36,30,33,33,29,31,28,26,33,33,37,32,31,34,37,30,35,28,35,35,33,39,32,32,38,29,28,30,33,35,31,34,33,40,36,35,36,34,36,32,36,30,31,33,35,37,33,32,31,39,39,34,37,31,33,36)
)
```

## Boxplot

```{r boxplot}
ggplot(bacteria_data, aes(x = pollutant, y = doubling_time, fill = pollutant)) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal(base_size = 14) +
  labs(title = "Bacterial Doubling Time by Pollutant Treatment",
       x = "Pollutant",
       y = "Doubling Time (minutes)") +
  theme(legend.position = "none")
```

## Summary Statistics

```{r summary}
bacteria_data %>%
  group_by(pollutant) %>%
  summarise(
    Mean = round(mean(doubling_time),2),
    SD   = round(sd(doubling_time),2),
    n    = n()
  )
```

## One-Way ANOVA

```{r anova}
bartlett.test(doubling_time ~ pollutant, data = bacteria_data)

model1 <- aov(doubling_time ~ pollutant, data = bacteria_data)
summary(model1)

TukeyHSD(model1)
```

# =====================================================
# PART 2 — ENZYME ACTIVITY (Two-Way ANOVA)
# =====================================================

```{r enzyme-data}
enzyme_data <- data.frame(
  temp = factor(rep(c("Low","Medium","High","Ultra High"), each = 50),
                levels = c("Low","Medium","High","Ultra High")),
  cellbatch = factor(rep(rep(LETTERS[1:5], each = 10), 4)),
  activity = c(
    69,80,88,87,85,100,105,104,83,89,102,89,79,96,96,84,96,100,85,96,
    81,98,75,79,106,96,91,92,85,87,95,75,62,102,81,92,90,82,86,103,
    74,73,95,76,80,92,87,94,100,89,
    186,192,233,213,191,208,162,174,168,161,
    156,185,142,225,197,168,125,184,216,105,
    151,164,116,188,242,166,158,197,168,168,
    228,251,152,202,106,217,78,179,120,234,
    191,217,266,225,209,190,216,197,207,177,
    145,141,144,177,150,168,152,163,146,141,
    164,147,163,130,160,145,140,177,147,161,
    163,148,167,149,126,167,143,159,140,177,
    166,140,148,156,147,156,178,141,162,157,
    153,150,173,146,149,165,127,128,163,152,
    99,34,117,66,57,113,178,193,53,127,
    4,141,64,38,153,71,115,62,131,8,
    6,204,57,115,79,108,28,143,167,79,
    54,148,88,87,186,71,132,98,-22,133,
    24,20,79,117,86,217,47,55,27,110
  )
)
```

## Interaction Plot

```{r interaction-plot}
ggplot(enzyme_data, aes(x = temp, y = activity,
                        color = cellbatch, group = cellbatch)) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  theme_minimal(base_size = 14) +
  labs(title = "Interaction Plot: Temperature vs Cell Batch",
       x = "Temperature",
       y = "Mean Activity",
       color = "Cell Batch")
```

## Two-Way ANOVA

```{r two-way-anova}
model2 <- aov(activity ~ temp * cellbatch, data = enzyme_data)
summary(model2)
```
# =====================================================
# PART 3 — MOUSE ACTIVITY (Two-Way ANOVA)
# =====================================================

## Create Data
```{r mouse-data}
mouse_data <- data.frame(
  time = factor(rep(c("morning","midday","afternoon"), each = 21),
                levels = c("morning","midday","afternoon")),
  
  light = factor(rep(rep(c("low","medium","bright"), each = 7), times = 3),
                 levels = c("low","medium","bright")),
  
  activity = c(
    14.1,5,14.7,7.3,8.7,5.3,13.3,
    14.5,10.8,11.4,11.7,8.6,10.1,16.8,
    10.1,9.6,11.7,19.2,21.8,13,2.7,
    
    16,18.7,9.1,8.9,11.9,10.6,18.8,
    6.6,7.8,14.3,10.3,12,21.9,17.9,
    13.9,18.6,8.3,21.1,28.2,11.5,14.5,
    
    12.5,16.1,9.3,9.3,12,12.4,9.4,
    9.1,20.8,16.7,13.7,17.6,11.9,4.8,
    20.6,19.4,12.4,14.9,14.2,9.5,19.7
  )
)

dim(mouse_data)
table(mouse_data$time, mouse_data$light)

model3 <- aov(activity ~ time * light, data = mouse_data)
summary(model3)

bartlett.test(activity ~ interaction(time, light), data = mouse_data)

shapiro.test(residuals(model3))

TukeyHSD(model3)

model_time_only    <- aov(activity ~ time, data = mouse_data)
model_light_only   <- aov(activity ~ light, data = mouse_data)
model_additive     <- aov(activity ~ time + light, data = mouse_data)
model_interaction  <- aov(activity ~ time * light, data = mouse_data)

AIC(model_time_only,
    model_light_only,
    model_additive,
    model_interaction)
